---
title: "experiment_1_cummulative_flux_analysis"
author: "KEY"
date: "2023-11-28"
output: html_document
---
  
This code corresponds to Figure 2 and Table S1 in the manuscript 
It's purpose is to graph and model the cumulative GPP of different biocrust
types (Type) given different experimental precipitation pulses (Treat)

```{r libraries, message=F}  
library(ggplot2)
library(tidyr)
library(tidyverse)
library(sjPlot)
library(car)
```

Read in the data

```{r }
data <- read_csv("experiment_1_flux_curve.csv") %>% 
  mutate(Treat = as.numeric(Treat), Type = as.factor(Type))  # Convert Treat to numeric and Type to factor
```

Exclude NAs and summarize GPP by Type, Treat, and Rep

```{r}
sum <- data %>%
  filter(!is.na(GPP_positive)) %>%  # Exclude rows with NA in GPP_positive
  group_by(Type, Treat, Rep) %>%
  summarize(sumGPP = sum(GPP_positive), .groups = 'drop')  # Sum GPP for each group
```

Model the data

```{r }
lm_model <- lm(sumGPP ~ Treat * Type, data = sum, na.action = na.exclude)
summary(lm_model)  # Check the summary of the model
tab_model(lm_model)
```

Check model assumptions

1. Normality of residuals

```{r }
residuals <- residuals(lm_model)
shapiro_test <- shapiro.test(residuals)  # Shapiro-Wilk test for normality
print(shapiro_test)
```

residuals are normally distributed according to Shapiro-Wilk test

Plot residuals

```{r }
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

residuals are roughly bell-shaped and do not show significant skew

2. Homogeneity of variance

Levene's test requires a formula that includes the grouping variable

```{r}
levene_test <- leveneTest(sumGPP ~ Type, data = sum)  # Test for homogeneity of variances across Types
print(levene_test)
```

Assumptions of homogeneity of variance hold

Plot diagnostic plots for ANOVA

```{r}
par(mfrow = c(2, 2))
plot(lm_model)
```

residuals vs fitted show no clear pattern
residuals fall along 45-degree angle
random spread of points
a few lines are outside of Cook's distance line, but not too abd

Plot data with linear model

```{r}
ggplot(sum, aes(x = Treat, y = sumGPP, color = Type, shape = Type)) + 
  geom_point(size = 3, position = position_jitter(width = 0.2, height = 0)) + # Points with jitter
  geom_smooth(method = "lm", aes(group = Type), se = TRUE) +  # Linear model with confidence interval
  ylab('Total measured GPP (µmol CO2/m²/s)') +
  xlab('Watering amount (mm)') +
  scale_color_manual(values = c("#fdbb84", "#2ca25f")) +  # Custom colors for Type
  scale_shape_manual(values = c("moss" = 17, "cyano" = 16)) +
  scale_x_continuous(breaks = c(1.8, 2.7, 5.4, 7.7, 10)) +  # Custom x-axis tick marks
  theme_bw() +
  theme(axis.title.x = element_text(size = 12),
        axis.text.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 16))
```
