---
title: "experiment_2_soil_moisture_analysis"
author: "KEY"
date: "2023-11-28"
output: html_document
---
  
This code corresponds to Figure S2 in the manuscript
The purpose of this is to model and graph the total time spent
wet (wetsum) for each biocrust type (crust_type) across each
experimental precipitation pulse treatment (treatment)
code to calculate total time spent wet can be found in 
"experiment_2_dry_down_curve_analysis"

```{r}
library(tidyr)
library(ggplot2)
library(tidyverse)
library(mgcv)
library(lme4)
library(sjPlot)
library(car)
```

reading in the data

```{r}
df <- read_csv("experiment_2_soil_moisture.csv") %>%
  mutate(treatment = as.factor(treatment), crust_type = as.factor(crust_type))
```

graphing the wetsum across treatments and crust types

```{r}
ggplot(df, aes(x = treatment, y = wetsum, fill = crust_type), size = 2) + 
  geom_boxplot() + 
  facet_grid(. ~ crust_type) +
  ylab('Culumative time spent wet (hours)') +
  xlab('Watering Treatment') +
  scale_x_discrete(breaks=c(1.8, 2.7, 5.4, 7.7, 10)) +
  scale_fill_manual(values = c("#fdbb84", "#2ca25f")) +
  theme_bw() +
  theme(axis.title.x=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.y=element_text(size=12),
        axis.text.y=element_text(size=12),
        panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.text.x=element_text(size=12),
        strip.text.y=element_text(size=12))
```

analyzing the variance of the total time spent wet across crust types

```{r}
t_test <- t.test(wetsum ~ crust_type, data = df)
print(t_test)
```

Get residuals from the t-test model
First calculate the difference between actual values and group means

```{r}
group_means <- tapply(df$wetsum, df$crust_type, mean)
residuals <- df$wetsum - group_means[df$crust_type]
```

Shapiro-Wilk test for normality

```{r}
shapiro_test <- shapiro.test(residuals)
print(shapiro_test)  # p-value > 0.05 suggests normality
```

this is not completely normal but is mostly normal

Visualize the residuals
Histogram of residuals

```{r}
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

Q-Q plot to check normality

```{r}
qqnorm(residuals)
qqline(residuals, col = "red")
```

the data mostly follows the 45-degree angle line

analyzing the variance of the total time spent wet across treatment types within cyano crusts

```{r}
cyano_df <- subset(df, crust_type == "cyano")
aov_cyano <- aov(wetsum ~ treatment, data = cyano_df)
summary(aov_cyano)
TukeyHSD(aov_cyano)
tab_model(aov_cyano)
```

Check model assumptions

1. Normality of residuals

```{r}
residuals <- residuals(aov_cyano)
shapiro_test <- shapiro.test(residuals)  # Shapiro-Wilk test for normality
print(shapiro_test)
```

residuals are distributed normally according to Shapiro-Wilk test

Plot residuals

```{r}
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

there is a bit of skew but linear models are robust to minor deviation

2. Homogeneity of variance
Levene's test requires a formula that includes the grouping variable

```{r}
levene_test <- leveneTest(wetsum ~ treatment, data = cyano_df)  # Test for homogeneity of variances across Types
print(levene_test)
```

Assumptions of homogeneity of variance are met

Plot diagnostic plots for ANOVA

```{r}
par(mfrow = c(2, 2))
plot(aov_cyano)
```

residuals vs fitted are clumped but the rest of model assumptions are met
residuals fall along 45-degree angle


analyzing the variance of the total time spent wet across treatment types within moss crusts

```{r}
moss_df <- subset(df, crust_type == "moss")
aov_moss <- aov(wetsum ~ treatment, data = moss_df)
summary(aov_moss)
tab_model(aov_moss)
```

Check model assumptions

1. Normality of residuals

```{r}
residuals <- residuals(aov_moss)
shapiro_test <- shapiro.test(residuals)  # Shapiro-Wilk test for normality
print(shapiro_test)
```

residuals are distributed normally according to Shapiro-Wilk test

Plot residuals

```{r}
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

there is a bit of skew but linear models are robust to minor deviation

2. Homogeneity of variance
Levene's test requires a formula that includes the grouping variable

```{r}
levene_test <- leveneTest(wetsum ~ treatment, data = moss_df)  # Test for homogeneity of variances across Types
print(levene_test)
```

Assumptions of homogeneity of variance are met

Plot diagnostic plots for ANOVA

```{r}
par(mfrow = c(2, 2))
plot(aov_moss)
```

residuals vs fitted are clumped but the rest of model assumptions are met
residuals fall along 45-degree angle
it doesn't look amazing, but it is ok for the purposes needed here

