---
title: "210_minus_60_GPP"
author: "RFH"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 201 minus 60 GPP

This R Markdown will track the analysis of GPP data collected by Young et al. First step will be to load the libraries and the data. 

```{r libraries, message=F}
library(ggplot2)
library(tidyr)
library(tidyverse)
library(mgcv)
library(lme4)
library(sjPlot)
library(car)
library(multcomp)
library(emmeans)
library(cowplot)
library(multcompView)
```

Because this is an R markdown file, the working directory is set to where ever this is saved. 

```{r datasets}
data <- read_csv("experiment_1_flux_curve.csv") %>% 
  mutate(Treat = as.numeric(Treat), Type = as.factor(Type))  # Convert Treat to numeric and Type to factor

data<-data %>% 
  mutate(Treat=as.factor(Treat), Type=as.factor(Type)) %>% #creates variables into factors for analysis
  mutate(Time = case_when(
    Time == 1 ~ 30,
    Time == 2 ~ 60,
    Time == 3 ~ 90,
    Time == 4 ~ 120,
    Time == 5 ~ 150,
    Time == 6 ~ 180,
    Time == 7 ~ 210))
  
```

## BUilding GAM for comparing crust type and water type

The objective of these analyses is to compare CO2 response rats of cyanobacteria and moss to different watering treatments (Treat). Original models were built by Kristin Young.

```{r original models}
model_1<- gam(GPP_positive ~ Type * Treat + s(Time, by = Treat, k = 2) +
               s(Rep, bs = 're'), 
             data = data, family=gaussian)

model_2<- gam(GPP_positive ~ Type * Treat + s(Time, by = Treat, k = 4) +s(Time, by = Type, k = 4) +
               s(Rep, bs = 're'), 
             data = data, family=gaussian) 

model_3<- gam(GPP_positive ~ Type * Treat +s(Time, by = Type, k = 4) +
               s(Rep, bs = 're'), 
             data = data, family=gaussian) 

AIC(model_1, model_2, model_3)## model 2 is lower AIC

summary(model_2)
plot(model_2)
anova(model_2)

tab_model(model_2)# R2=0.664, can use this output in MS to explain model observations
gam.check(model_2) #model looks really good with assumptions

#Anova(model_2)
```

The model looks pretty good and is explaining about 70% of the variation in the data. To help with interpretation, we can now build out predictive models so we can plot our model output and see how it looks.

## Build models and think about stats tables

```{r model prediction}


predict<-tidymv::predict_gam(model_2) 

#test plot of the curves from the model
predict %>% 
  ggplot(aes(x = Time, y = fit, color = Type))+
  geom_smooth(method="gam", formula=y~s(x))+
  facet_grid(. ~ Treat, scales = 'free') 
```
Trying to figure out how to remove the gaps in the confidence intervals.

Next step is to clean up the plot and bring it all together.
```{r Figure 1}
graph_df<-predict%>% #first need to merge in the predictive data
   mutate(lwr=fit-(2*se.fit), upr=fit+(2*se.fit))

(fig1<-
  ggplot()+
    geom_ribbon(data=graph_df, aes(x=Time, ymin=lwr, ymax= upr, group=Type),  fill="gray")+
  stat_smooth(data=graph_df, aes(x=Time, y=fit, color=Type), method="gam", formula=y~s(x), se=T)+
    geom_point(data=data,aes(x=Time, y= GPP_positive, color = Type, shape= Type))+
   facet_grid(. ~ Treat, scales = 'free') + 
  ylab('GPP (µmol CO2/m²/s)') +
  xlab('Time (minutes)') +
  scale_color_manual( "Crust Type", values = c("#fdbb84", "#2ca25f")) +
    scale_shape_manual( "Crust Type", values = c(16,17)) +
  scale_x_continuous(breaks=seq(0, 210, 30))+
  theme_bw() +
  theme(axis.title.x=element_text(size=12),
       axis.text.x = element_text(angle = 30, hjust = 1, size=9, color = "black"),
        axis.title.y=element_text(size=12),
        axis.text.y=element_text(size=12),
        panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12))
 )
fig1

ggsave("fig1.png", dpi=300, width=8, height=5.5, units="in")

```
Figure out the stats for this above figure and make a table. Are the patterns different between moss and cyano overtime within each treatment? I think the best option is to fit a linear mixed effects model and then do post-hoc testing with the multcomp package.

```{r figure out post-hoc statistics}

model_2<- gam(GPP_positive ~ Type * Treat + s(Time, by = Treat, k = 4) +
                s(Time, by = Type, k = 4) +
               s(Rep, bs = 're'), 
             data = data, family=gaussian) 

mod_posthoc<-lme4::lmer(GPP_positive~Type*Treat+(1|Rep), data=data)

summary(mod_posthoc)
Anova(mod_posthoc)

emm = emmeans(mod_posthoc, ~ Type*Treat)
pairs(emm)

model_means_cld <- cld(object = emm,
                       adjust = "Tukey",
                       Letters = letters,
                       alpha = 0.05)

model_means_cld

```


