---
title: "experiment_2_initial_final_flux"
author: "KEY"
date: "2023-11-28"
output: html_document
---
 
This code corresponds to Figure 3 and Table S2 in the manuscript
The purpose of this code is to graph and analyze CO2 flux data (NSE, GPP, Respiration)
from different biocrust types (Type) given different precipitation pulse 
treatments (Treat) at the beginning of the experiment (Initial) and the
end of the experiment (Final)

```{r}
library(tidyr)
library(ggplot2)
library(tidyverse)
library(mgcv)
library(lme4)
library(sjPlot)
library(car)
```

reading in and assigning flux data

```{r}
flux_data <- read_csv("experiment_2_initial_final_flux.csv") %>% 
  mutate(Treat = as.numeric(Treat), Type = as.factor(Type), GPP_positive = as.numeric(GPP_positive)) 

# flux_data <- read_csv("experiment_2_initial_final_flux.csv") %>% 
 # mutate(Treat = as.numeric(Treat), Type = as.factor(Type), GPP_positive = as.numeric(GPP_positive)) #creates variables into factors and numeric for analysis
```

renaming flux values

```{r}
real <- rename(flux_data, GPP = "GPP_positive",
              NSE = "NSE_Lin_Flux_Light",
              Respiration = "R_Dark_Flux_Light")
```

renaming timmepoin values

```{r}
real$Timepoint <- factor(real$Timepoint, levels = c("Initial", "Final"),
                  labels = c("Initial Treatment", "Final Treatment")
)
```

grouping data to show in a single graph

```{r}
real_combined <- real %>%
  gather(Flux, Value, c("GPP", "Respiration", "NSE")) %>%
  mutate(Value = as.numeric(Value))
```

graphing the initial vs final flux data for GPP, NSE, Respiration

```{r}
ggplot(real_combined, aes(x = Treat, y = Value, color = Type, shape = Type), size = 2) + 
  geom_smooth(method = lm) + 
  facet_grid(Flux ~ Timepoint) +
  geom_point(position = position_dodge(width = 0.5)) +
  ylab('µmol CO2 m-2s-1') +
  xlab('Watering amount (mm)') +
  scale_x_continuous(breaks=c(1.8, 2.7, 5.4, 7.7, 10)) +
  scale_color_manual(values = c("#fdbb84", "#2ca25f")) +
  theme_bw() +
  theme(axis.title.x=element_text(size=12),
        axis.text.x=element_text(size=12),
        axis.title.y=element_text(size=12),
        axis.text.y=element_text(size=12),
        panel.grid.major = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.text.x=element_text(size=12),
        strip.text.y=element_text(size=12))
```

INITIAL TIMEPOINT ANALYSIS

```{r}
lm_pre <- subset(real, Timepoint == "Initial Treatment")
lm_pre$Treat <- as.numeric(lm_pre$Treat)
lm_pre$GPP <- as.numeric(lm_pre$GPP)
```

INITIAL GPP
making a linear model for GPP at the initial timepoint

```{r}
lm_pre_GPP_model <- lm(GPP ~ Type * Treat, data = lm_pre)
summary(lm_pre_GPP_model)
tab_model(lm_pre_GPP_model)
```

Check model assumptions

1. Normality of residuals

```{r}
residuals <- residuals(lm_pre_GPP_model)
shapiro_test <- shapiro.test(residuals)  # Shapiro-Wilk test for normality
print(shapiro_test)
```

residuals are normally distributed according to Shapiro-Wilk test

Plot residuals

```{r}
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

residuals are roughly bell-shaped and do not show significant skew

2. Homogeneity of variance
Levene's test requires a formula that includes the grouping variable

```{r}
levene_test <- leveneTest(Respiration ~ Type, data = lm_pre)  # Test for homogeneity of variances across Types
print(levene_test)
```

Assumptions of homogeneity of variance hold

Plot diagnostic plots for ANOVA

```{r}
par(mfrow = c(2, 2))
plot(lm_pre_GPP_model)
```

residuals vs fitted show no clear pattern
residuals fall along 45-degree angle
random spread of points
a few lines are outside of Cook's distance line, but not too bad

INITIAL NSE
making a linear model for NSE at the initial timepoint

```{r}
lm_pre_NSE_model <- lm(NSE ~ Type * Treat, data = lm_pre)
summary(lm_pre_NSE_model)
tab_model(lm_pre_NSE_model)
```


Check model assumptions

1. Normality of residuals

```{r}
residuals <- residuals(lm_pre_NSE_model)
shapiro_test <- shapiro.test(residuals)  # Shapiro-Wilk test for normality
print(shapiro_test)
```

residuals are normally distributed according to Shapiro-Wilk test

Plot residuals

```{r}
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

residuals are roughly bell-shaped and do not show significant skew

2. Homogeneity of variance
Levene's test requires a formula that includes the grouping variable

```{r}
levene_test <- leveneTest(NSE ~ Type, data = lm_pre)  # Test for homogeneity of variances across Types
print(levene_test)
```

Assumptions of homogeneity of variance hold

Plot diagnostic plots for ANOVA

```{r}
par(mfrow = c(2, 2))
plot(lm_pre_NSE_model)
```

residuals vs fitted show no clear pattern
residuals fall along 45-degree angle
random spread of points
a few lines are outside of Cook's distance line, but not too bad

INITIAL RESPIRATION
making a linear model for R at the initial timepoint

```{r}
lm_pre_R_model <- lm(Respiration ~ Type * Treat, data = lm_pre)
summary(lm_pre_R_model)
tab_model(lm_pre_R_model)
```

Check model assumptions

1. Normality of residuals

```{r}
residuals <- residuals(lm_pre_R_model)
shapiro_test <- shapiro.test(residuals)  # Shapiro-Wilk test for normality
print(shapiro_test)
```

residuals are normally distributed according to Shapiro-Wilk test

Plot residuals

```{r}
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

residuals are roughly bell-shaped and do not show significant skew

2. Homogeneity of variance
Levene's test requires a formula that includes the grouping variable

```{r}
levene_test <- leveneTest(Respiration ~ Type, data = lm_pre)  # Test for homogeneity of variances across Types
print(levene_test)
```

Assumptions of homogeneity of variance hold

Plot diagnostic plots for ANOVA

```{r}
par(mfrow = c(2, 2))
plot(lm_pre_R_model)
```

residuals vs fitted show no clear pattern
residuals fall along 45-degree angle
random spread of points
a few lines are outside of Cook's distance line, but not too bad

FINAL TREATMENT ANALYSIS

```{r}
lm_post <- subset(real, Timepoint == "Final Treatment")
lm_post$Treat <- as.numeric(lm_post$Treat)
```

FINAL GPP
making a linear model for GPP at the final timepoint

```{r}
lm_post_GPP_model <- lm(GPP ~ Type * Treat, data = lm_post)
summary(lm_post_GPP_model)
tab_model(lm_post_GPP_model)
```

Check model assumptions

1. Normality of residuals

```{r}
residuals <- residuals(lm_post_GPP_model)
shapiro_test <- shapiro.test(residuals)  # Shapiro-Wilk test for normality
print(shapiro_test)
```

residuals are mostly normally distributed according to Shapiro-Wilk test

Plot residuals

```{r}
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

there is a bit of skew but linear models are robust to minor deviation

2. Homogeneity of variance

Levene's test requires a formula that includes the grouping variable

```{r}
levene_test <- leveneTest(GPP ~ Type, data = lm_post)  # Test for homogeneity of variances across Types
print(levene_test)
```

Assumptions of homogeneity of variance hold

Plot diagnostic plots for ANOVA

```{r}
par(mfrow = c(2, 2))
plot(lm_post_GPP_model)
```

residuals vs fitted show no clear pattern
residuals fall along 45-degree angle
random spread of points
a few lines are outside of Cook's distance line, but not too bad

FINAL NSE
making a linear model for NSE at the final timepoint

```{r}
lm_post_NSE_model <- lm(NSE ~ Type * Treat, data = lm_post)
summary(lm_post_NSE_model)
tab_model(lm_post_NSE_model)
```

Check model assumptions
1. Normality of residuals

```{r}
residuals <- residuals(lm_post_NSE_model)
shapiro_test <- shapiro.test(residuals)  # Shapiro-Wilk test for normality
print(shapiro_test)
```

residuals are not normally distributed according to shapiro-wilks

Plot residuals

```{r}
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

It looks like there are two data points skewing the data, the model is robust 
enough to deal with two skewed points

2. Homogeneity of variance
Levene's test requires a formula that includes the grouping variable

```{r}
levene_test <- leveneTest(NSE ~ Type, data = lm_post)  # Test for homogeneity of variances across Types
print(levene_test)
```

Assumptions of homogeneity of variance hold

Plot diagnostic plots for ANOVA

```{r}
par(mfrow = c(2, 2))
plot(lm_post_NSE_model)
```

residuals vs fitted show no clear pattern
residuals fall along 45-degree angle
random spread of points
a few lines are outside of Cook's distance line, but not too bad

FINAL RESPIRATION
making a linear model for R at the final timepoint

```{r}
lm_post_R_model <- lm(Respiration ~ Type * Treat, data = lm_post)
summary(lm_post_R_model)
tab_model(lm_post_R_model)
```

Check model assumptions

1. Normality of residuals

```{r}
residuals <- residuals(lm_post_R_model)
shapiro_test <- shapiro.test(residuals)  # Shapiro-Wilk test for normality
print(shapiro_test)
```

residuals are mostly normally distributed according to Shapiro-Wilk test

Plot residuals

```{r}
ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Residuals", x = "Residuals", y = "Frequency")
```

there is a bit of skew but linear models are robust to minor deviation

2. Homogeneity of variance

Levene's test requires a formula that includes the grouping variable

```{r}
levene_test <- leveneTest(Respiration ~ Type, data = lm_post)  # Test for homogeneity of variances across Types
print(levene_test)
```

Assumptions of homogeneity of variance are not met, but its not too bad

Plot diagnostic plots for ANOVA

```{r}
par(mfrow = c(2, 2))
plot(lm_post_R_model)
```

residuals vs fitted show no clear pattern
residuals fall along 45-degree angle
random spread of points
a few lines are outside of Cook's distance line, but not too bad

